### 2010-04-30 VerboseNotes.tex
CheatSheet Update
git-svn-id: https://dav.misho-web.com/svn/cyghome/trunk/QFT/CheatSheet@386 c04c31ef-c4f7-4b37-ba94-86ebe2b9c820

%#!latexmk CheatSheet.tex
%%% Time-Stamp: <2010-04-29 16:59:55 misho>
\section{Verbose Notes}
\subsection{Spinor Fields}
\def\redblue#1#2{({\RED{#1}},{\BLUE{#2}})}

\subsubsection{Lorentz group and Lorentz algebra}
\begin{tabular}{l@{ :\ \ \ }l}
Metric &  ${\RED\eta} = \diag(+1, -1, -1, -1),\qquad {\BLUE \eta} = \diag(-1,+1,+1,+1).$\\
Lorentz transf. in $\mathbb R^{1,3}$ & Linear transf. $x^\mu \mapsto \Lambda\T^\mu_\nu x^\nu$ which conserve $x^2$.\\
 & \then $\Hrsd = \Hmnd \Lambda\T^\mu_\rho \Lambda\T^\nu_\sigma.$
 and form a group $L$.\\
 & \qquad
   ($\then\quad
     (\Lambda^{-1})\T^\mu_\nu=\eta_{\nu\alpha}\eta^{\mu\beta}\Lambda\T^\alpha_\beta=:\Lambda\T_\nu^\mu\quad\then\quad\Lambda\T^\mu_\nu\Lambda\T_\mu^\rho=\delta^\rho_\nu$)\\[.8zw]

Disconnected parts of $L$ &
  $L_0 := \left\{ \det\Lambda = +1 \land \Lambda\T_0^0 >0\right\} \quad
   L_P := \left\{ \det\Lambda = -1 \land \Lambda\T_0^0 >0\right\}$\\
& $L_T := \left\{ \det\Lambda = +1 \land \Lambda\T_0^0 <0\right\} \quad
   L_{PT} := \left\{ \det\Lambda = -1 \land \Lambda\T_0^0 <0\right\}$\\
& \qquad ($L_0$ is identical with $\redblue{\gSO(1,3)}{\gSO(3,1)}$. )\\[1zw]
Infinitesimal one in $L_0$& $\Lambda\T^\mu_\nu = \delta^\mu_\nu + \epsilon\T^\mu_\nu$ where $\epsilon_{\mu\nu}=-\epsilon_{\nu\mu}$ (for $\eta=\eta\Lambda\Lambda$)\\[1zw]
\end{tabular}

\vspace{1zw}

微小変換は$\epsilon\T^\mu_\nu=\spmat{
0&\beta_x&\beta_y&\beta_z\\
\beta_x & 0 & -\theta_z & \theta_y\\
\beta_y & \theta_z & 0 & -\theta_x\\
\beta_y & -\theta_y & \theta_x & 0}$の形となっているので，回転生成子$\vc J$と加速生成子$\vc K$は
\begin{equation}
 \Lambda = \exp\left[\kappa\ii(\vipro\theta J +\vipro\beta K)\right]
\quad\then\quad
J_x = -\kappa\ii\spmat{0&0&0&0\\0&0&0&0\\0&0&0&-1\\0&0&1&0},\quad
K_x = -\kappa\ii\spmat{0&1&0&0\\1&0&0&0\\0&0&0&0\\0&0&0&0}\label{eq:explicit_JK}
\end{equation}
の形である。ここで$\kappa=\pm1$はnotationである。

一方，微小変換から生成子を
 $\epsilon\T^\mu_\nu  =: \mp\frac\ii2\epsilon^{\rho\sigma}(J_{\rho\sigma})\T^\mu_\nu$
と定義すると，計量によらずに$\epsilon_{\mu\nu}$は反対称となり，従って$J^{\rho\sigma}$も反対称。
$ (J_{\rho\sigma})\T^\mu_\nu =
\pm\ii\left(\delta^\mu_\rho\Hsnd-\delta^\mu_\sigma\Hrnd\right)$
となり，交換関係が得られ，これが閉じているのでLie代数であることもわかる：
\begin{equation}
 [J_{\mu\nu},J_{\rho\sigma}] = \mp\ii(
    \Hmrd J_{\nu\sigma} + \Hnsd J_{\mu\rho} - \Hmsd J_{\nu\rho} - \Hnrd J_{\mu\sigma}).
\end{equation}

生成子の具体形は計量に依存し，
$
 \displaystyle J\T_1_0^\mu_\nu=\redblue{\pm\ii}{\mp\ii}
 \spmat{0&1&0&0\\1&0&0&0\\0&0&0&0\\0&0&0&0}_{\mu\nu},
 \displaystyle J\T_2_3^\mu_\nu=\redblue{\pm\ii}{\mp\ii}
 \spmat{0&0&0&0\\0&0&0&0\\0&0&0&-1\\0&0&1&0}_{\mu\nu}
$となる。
よって，ここでの複号の取り方と$\kappa$および計量の定義によって，$\vc J$・$\vc K$と$J_{\rho\sigma}$の対応が定まることになる。
\begin{equation}
  \vc\theta = \redblue+-(\epsilon^{23},\epsilon^{31},\epsilon^{12}),\quad
  \vc\beta  = \redblue+-(\epsilon^{10},\epsilon^{20},\epsilon^{30}).
\end{equation}
\starline

以下では$\kappa=-1$にとり，複号の$({\RED 上},{\BLUE 下})$側を取ることで
\begin{equation}
  \vc J     = (J_{23}, J_{31}, J_{12}),\quad  \vc K = (J_{10},J_{20},J_{30});\qquad
\end{equation}
とする：
\begin{equation}
 \Lambda = \exp\epsilon = \exp\left[-\ii(\vipro\theta J +\vipro\beta K)\right]
         = \exp\left[\redblue{-\ii}{+\ii}(\epsilon^{\rho\sigma}J_{\rho\sigma})/2\right].
\end{equation}

\newpage
\subsubsection{Lorentz group and $\gSL(2,\Complex)$}
次に，連結Lie群$L_0$が，連結Lie群$\gSL(2,\Complex)/\gZ_2$と同型であることを見る：
\begin{align}
 \aSL(2,\Complex) :=& \{a\in \aGL(2,\Complex)\ |\ \Tr(a)=0\},&
 \gSL(2,\Complex) :=& \{g\in \gGL(2,\Complex) |\ \det(g)=1\}.
\end{align}
まず,$\Sm$を（極めて一般的に） $\Sm:=(\alpha1,\beta\vc\sigma)$と定義する($\alpha=\beta=\pm1$)。
$x^2 = ({\RED+}{\BLUE-})\det(x_\mu\Sm)$なので
\begin{equation}
 f^g: (x_\mu\Sm)\mapsto g (x_\mu\Sm) g^\dagger;\quad g\in\gSL(2,\Complex)
\end{equation}
は$x^2$を保存する。よってLorentz変換であり，生成子を比べることで局所同型だとわかる：
\begin{rightnote}
$\gSL(2,\Complex)\ni g = \exp(-\ii a)$として
$x_\mu (g \Sm g^\dagger) = \Lambda\T_\mu^\nu x_\nu \Sm$を微小展開すると
\begin{equation}
\Lambda\T^\mu_\nu\Sn=g^{-1}\Sm (g^{-1})^\dagger\quad\then\quad
\mp\frac12\epsilon_{\rho\sigma}(J^{\rho\sigma})\T^\mu_\nu\Sn=a\Sm-\Sm a^\dagger
\end{equation}
\TODO{正しく複号をとって再チェック！}

$({\RED\epsilon^{10}},{\BLUE\epsilon^{01}})=1$のとき
$a = -\dfrac\ii2\alpha\beta\sigma_x$
であり，$({\RED\epsilon^{23}},{\BLUE\epsilon^{32}})=1$のとき
$a = \dfrac12\sigma_x$である。
\end{rightnote}

このことを別の観点から見る。Lorentz群の生成子の交換関係を見ると，（正しく複号を取った場合
\begin{equation}
  [J_i, J_j]= \ii\epsilon_{ijk}J_k,\qquad
 [J_i, K_j]= \ii\epsilon_{ijk}K_k,\qquad
 [K_i, K_j]= -\ii\epsilon_{ijk}J_k
\end{equation}
となるので，
\begin{equation}
  \vc A:=\frac{1}2(\vc J+\ii\vc K), \qquad \vc B := \frac{1}2(\vc J-\ii\vc K).
\end{equation}
と定義すると
\begin{equation}
  [A_i, A_j] = \ii\epsilon_{ijk}A_k,\qquad
 [B_i, B_j] = \ii\epsilon_{ijk}B_k,\qquad
 [A_i, B_j] = 0,
\end{equation}
となり，Lorentz群が$\gSU(2)\times\gSU(2)$に分解できる。

\subsection{Weyl Spinor for $\eta=(+,-,-,-)$}
左巻きspinor $\xi$ と右巻きspinor $\bar\xi$ を，
\begin{align}
 \xi&\mapsto\left(1-\frac\ii2\vipro\theta\sigma-\frac12\vipro\beta\sigma\right)\xi&
 \bar\xi&\mapsto\left(1-\frac\ii2\vipro\theta\sigma+\frac12\vipro\beta\sigma\right)\bar\xi
\end{align}
と変換するものとして定義する。
$\gSU(2)_A\times\gSU(2)_B$に対して左巻きは$(1/2,0)$表現，右巻きは$(0,1/2)$表現となっている。

$g$を左巻きspinorの変換子として定義すると，$\vc J\to\frac12\vc \sigma$および$\vc K\to-\frac\ii2\alpha\beta\vc\sigma$の対応があったことを思い出すと
$-\ii(\vipro\theta J+\vipro\beta K)=-\frac\ii2\vipro\theta\sigma-\frac12\alpha\beta\vipro\beta\sigma$であるので，$\alpha\beta=1$が分かる。また，記号を
\begin{align}
  \xi_\alpha      &\mapsto g\T_\alpha^\beta\xi_\beta,
& \bar\xi^\dalpha &\mapsto (g^\dagger)^{-1}\T^\dalpha_\dbeta\bar\xi^\dbeta
\end{align}
と定義しておく。

次に，$\xi^\alpha\chi_\alpha$および$\bar\xi_\dalpha\bar\chi^\dalpha$がscalarとなるようにしたい。
$E=\spmat{0&1\\-1&0}$としておくと，$-E\trans{g}E=g^{-1}$であるので，
\begin{equation}
 (\xi')^\alpha=\xi^\beta(g^{-1})\T_\beta^\alpha
              =-\xi^\beta(E_{\beta A}g\T_B^AE_{B\alpha})\qquad\therefore (-E_{\gamma\alpha})(\xi')^\alpha = -g\T_\gamma^\beta(E_{\beta\delta}\xi^\delta).
\end{equation}
よって，$\epsilon^{12}=\epsilon_{21}=1$として
$\xi^\alpha:=\epsilon^{\alpha\beta}\xi_\beta$，$\xi_\alpha=\epsilon_{\alpha\beta}\xi^\beta$
とすれば良い。

同様に，$\bar\xi'_\dalpha=-E(g^\dagger)^{-1}E\bar\xi_\dbeta$から
$(E^{\dalpha\dbeta}\bar\xi_\dbeta)'=(g^\dagger)^{-1}{}\T^\dalpha_\dbeta (E^{\dbeta\dgamma}\bar\xi_\dgamma)$となる。$\epsilon_{12}=\epsilon_{\dot1\dot2}$として$\bar\xi_\dalpha:=\epsilon_{\dalpha\dbeta}\bar\xi^\dalpha$とするのが一般的である。
また，このことから$(\xi_a)^*=\bar\xi_\dalpha$（或いは$\xi^\dagger=\bar\xi$）が分かる。

$x_\mu\Sm\mapsto x_\mu(g\Sm g^\dagger)$であるので，
$x_\mu(\xi^\alpha \Sm \bar\chi^\dalpha)$はscalarである。よって，$\Sm_{\alpha\dalpha}$のように書ける。
更にここで$x_\mu(\bar\xi_\dalpha\bSm{}^{\dalpha\alpha}\chi_\alpha)$もscalarとなるように$\bar\sigma$を定めよう。
\begin{equation}
 \bar\xi_\dalpha\bSm{}^{\dalpha\alpha}\chi_\alpha=-\epsilon_{\alpha\beta}\epsilon_{\dalpha\dbeta}\chi^\beta\bSm{}^{\dalpha\alpha}\bar\xi^\dbeta
\qquad\therefore \Sm_{\beta\dbeta}\propto \epsilon_{\alpha\beta}\epsilon_{\dalpha\dbeta}\bSm{}^{\dalpha\alpha}
\end{equation}
であり，あとはconventionである。

\starline

\begin{align}
  \epsilon^{12}=\epsilon_{21}=\epsilon^{\dot1\dot2}=\epsilon_{\dot2\dot1}&=1,&
\xi^\alpha&:=\epsilon^{\alpha\beta}\xi_\beta,& \xi_\alpha&=\epsilon_{\alpha\beta}\xi^\beta,&
\bar\xi^\dalpha&:=\epsilon^{\dalpha\dbeta}\bar\xi_\dbeta,& \bar\xi_\dalpha&=\epsilon_{\dalpha\dbeta}\bar\xi^\dbeta
\end{align}
\begin{align}
  \xi_\alpha      &\mapsto g\T_\alpha^\beta\xi_\beta,
& \xi^\alpha      &\mapsto \xi^\beta (g^{-1})\T_\beta^\alpha,
& \bar\xi_\dalpha &\mapsto \bar\xi_\dbeta(g^\dagger)\T^\dbeta_\dalpha
& \bar\xi^\dalpha &\mapsto (g^\dagger)^{-1}\T^\dalpha_\dbeta\bar\xi^\dbeta
\end{align}
\begin{align}
  \bSm{}^{\dalpha\alpha} &:= \epsilon^{\alpha\beta}\epsilon^{\dalpha\dbeta}\Sm_{\beta\dbeta}&
 \Sm_{\alpha\dalpha}&= \epsilon_{\alpha\beta}\epsilon_{\dalpha\dbeta}\bSm{}^{\dbeta\beta}&
 \therefore \Sm:=(1,\vc\sigma),\quad\bSm:=(1,-\vc\sigma)
\end{align}

\subsection{Weyl Spinor for $\eta=(-,+,+,+)$}
左巻き／右巻きの定義は同じであり，それぞれやはり$(1/2,0)$表現と$(0,1/2)$表現である：
\begin{align}
 \xi&\mapsto\left(1-\frac\ii2\vipro\theta\sigma-\frac12\vipro\beta\sigma\right)\xi&
 \bar\xi&\mapsto\left(1-\frac\ii2\vipro\theta\sigma+\frac12\vipro\beta\sigma\right)\bar\xi
\end{align}

$g$を左巻きspinorの変換子として定義すると，$\vc J\to\frac12\vc \sigma$および$\vc K\to-\frac\ii2\alpha\beta\vc\sigma$の対応があったことを思い出すと
$-\ii(\vipro\theta J+\vipro\beta K)=\frac\ii2\vipro\theta\sigma+\frac12\alpha\beta\vipro\beta\sigma$であるので，$\alpha\beta=1$が分かる。
\begin{align}
  \xi_\alpha      &\mapsto g\T_\alpha^\beta\xi_\beta,
& \bar\xi_\dalpha &\mapsto \bar\xi_\dbeta(g^\dagger)\T^\dbeta_\dalpha
\end{align}
であり，$\bar\xi=\xi^\dagger$，つまり$\bar\xi_\dalpha = (\xi_\alpha)^*$である。

次に，$\xi^\alpha\chi_\alpha$および$\bar\xi_\dalpha\bar\chi^\dalpha$がscalarとなるようにしたい。
$E=\spmat{0&1\\-1&0}$としておくと，$-E\trans{g}E=g^{-1}$であるので，
\begin{equation}
 (\xi')^\alpha=\xi^\beta(g^{-1})\T_\beta^\alpha
              =-\xi^\beta(E_{\beta A}g\T_B^AE_{B\alpha})\qquad\therefore (-E_{\gamma\alpha})(\xi')^\alpha = -g\T_\gamma^\beta(E_{\beta\delta}\xi^\delta).
\end{equation}
よって，$\epsilon^{12}=\epsilon_{21}=1$として
$\xi^\alpha:=\epsilon^{\alpha\beta}\xi_\beta$，$\xi_\alpha=\epsilon_{\alpha\beta}\xi^\beta$
とすれば良い。

同様に，$(\bar\xi^\dalpha E_{\dalpha\dbeta})'=(\bar\xi^\dalpha E_{\dalpha\dbeta})'g^\dagger\T^\dgamma_\dbeta$……としても良かったが，左からの演算，および$\epsilon_{12}=\epsilon_{\dot1\dot2}$の性質を重視して$(\epsilon_{\dalpha\dbeta}\bar\xi^\dbeta)'=(\epsilon_{\dgamma\dbeta}\bar\xi^\dbeta)'g^\dagger\T^\dgamma_\dbeta$とするのが一般的である。

$x_\mu\Sm\mapsto x_\mu(g\Sm g^\dagger)$であるので，
$x_\mu(\xi^\alpha \Sm \bar\chi^\dalpha)$はscalarである。よって，$\Sm_{\alpha\dalpha}$のように書ける。
更にここで$x_\mu(\bar\xi_\dalpha\bSm{}^{\dalpha\alpha}\chi_\alpha)$もscalarとなるように$\bar\sigma$を定めよう。
\begin{equation}
 \bar\xi_\dalpha\bSm{}^{\dalpha\alpha}\chi_\alpha=-\epsilon_{\alpha\beta}\epsilon_{\dalpha\dbeta}\chi^\beta\bSm{}^{\dalpha\alpha}\bar\xi^\dbeta
\qquad\therefore \Sm_{\beta\dbeta}\propto \epsilon_{\alpha\beta}\epsilon_{\dalpha\dbeta}\bSm{}^{\dalpha\alpha}
\end{equation}
であり，あとはconventionである。

\starline

\begin{align}
  \epsilon^{12}=\epsilon_{21}=\epsilon^{\dot1\dot2}=\epsilon_{\dot2\dot1}&=1,&
\xi^\alpha&:=\epsilon^{\alpha\beta}\xi_\beta,& \xi_\alpha&=\epsilon_{\alpha\beta}\xi^\beta,&
\bar\xi^\dalpha&:=\epsilon^{\dalpha\dbeta}\bar\xi_\dbeta,& \bar\xi_\dalpha&=\epsilon_{\dalpha\dbeta}\bar\xi^\dbeta
\end{align}
\begin{align}
  \xi_\alpha      &\mapsto g\T_\alpha^\beta\xi_\beta,
& \xi^\alpha      &\mapsto \xi^\beta (g^{-1})\T_\beta^\alpha,
& \bar\xi_\dalpha &\mapsto \bar\xi_\dbeta(g^\dagger)\T^\dbeta_\dalpha
& \bar\xi^\dalpha &\mapsto (g^\dagger)^{-1}\T^\dalpha_\dbeta\bar\xi^\dbeta
\end{align}
\begin{align}
  \bSm{}^{\dalpha\alpha} &:= \epsilon^{\alpha\beta}\epsilon^{\dalpha\dbeta}\Sm_{\beta\dbeta}&
 \Sm_{\alpha\dalpha}&= \epsilon_{\alpha\beta}\epsilon_{\dalpha\dbeta}\bSm{}^{\dbeta\beta}&
 \therefore \Sm:=(1,\vc\sigma),\quad\bSm:=(1,-\vc\sigma)
\end{align}

\subsection{Polarization Sum}\label{Sec:Verbose:PolarizationSum}
Firstly we focus on the single photon case $M=\epsilon_\mu^*(k)\epsilon_\nu'^*(k')M^{\mu\nu}$.
Here we set $k=(E,0,0,E)$, and $\epsilon=(0,1,0,0)\oplus(0,0,1,0)$. Then
\begin{equation}
  \sum\s{pol.}\left|M\right|^2
= \sum\s{pol.}\epsilon_\mu^*(k)\epsilon_\nu(k)M^{\mu}M^{\nu*}
= |M^1|^2+|M^2|^2,
\end{equation}
while
\begin{equation}
 \Hmnd M^{\mu}M^{\nu*} = |M^1|^2+|M^2|^2
\end{equation}
for Ward identity $k_\mu M^\mu=0$. Therefore the replacement
\begin{equation}
 \sum\s{pol.}\epsilon_\mu\epsilon'_\nu \to \Hmnd
\end{equation}
is valid.

Secondly we think about the double photons case
\footnote{This part is derived from 濱口幸一's notebook.}
$M=\epsilon_\mu^*(k)\epsilon_\nu'^*(k')M^{\mu\nu}$.
Here we set
\begin{align}
 k &=(E,0,0, E) & \epsilon &=(0,1,0,0)\oplus(0,0,1,0)\\
 k'&=(E,0,0,-E) & \epsilon'&=(0,\cos\theta,\sin\theta,0)\oplus(0,-\sin\theta,\cos\theta,0).
\end{align}
Then doing some simple calculations, we can get
\begin{equation}
  \sum\s{pol.}\left|M\right|^2 = |M^{11}|^2+|M^{12}|^2+|M^{21}|^2+|M^{22}|^2.
\end{equation}
Nevertheless, na\"ive replacement does not work, because our Ward identities
\begin{equation}
 k_\mu\epsilon_\nu'^*(k')M^{\mu\nu} =
 \epsilon_\mu^*(k)k'_\nu M^{\mu\nu} = 0 \label{eq:WardIdentities}
\end{equation}
obviously does not help us. If we can omit $\epsilon$s from these
identities, that is if
\begin{equation}
 k_\mu M^{\mu\nu} = k'_\nu M^{\mu\nu} = 0,\label{eq:TrueIdentities}
\end{equation}
we can recover validity of the replacement:
\begin{align}
 \Hmrd\Hnsd M^{\mu\nu}M^{\rho\sigma*}
&= -\Hnsd\left(
M^{1\nu}M^{1\sigma*}+M^{2\nu}M^{2\sigma*}
\right)\\
&= |M^{11}|^2+|M^{12}|^2+|M^{21}|^2+|M^{22}|^2.
\end{align}

Then what's happening? Why this replacement is not valid?
Actually our new conditions \eqref{eq:TrueIdentities} seem to guarantee
that we are summing not only ``physical'' but also ``unphysical'' polarizations.
Meanwhile if we use some physical condition such as $\epsilon\cdot k=0$,
\eqref{eq:TrueIdentities} break down while Ward identities
\eqref{eq:WardIdentities} are still valid.

Now let's check what is happening from another viewpoint. First we suppose
$M$ satisfies our new conditions \eqref{eq:TrueIdentities}, and define
$\tilde M^{\mu\nu}$ and $\tilde M$ as
\begin{align}
 \tilde M^{\mu\nu}& := M^{\mu\nu} + k^\mu p^\nu + p'^\mu k'^\nu,\\
 \tilde M         & := \epsilon_\mu^*(k)\epsilon_\nu'^*(k')\tilde M^{\mu\nu}.
\end{align}
This alternative amplitude satisfies Ward identities (since photon is
massless and $\epsilon\cdot k=0$), and furthermore $\tilde M = M$.
Therefore $\tilde M$ is physically identical to $M$.
However technically these are very different, just because we cannot
perform our ``na\"ive replacement'' for this $\tilde M$:
\begin{align}
 \Hmrd\Hnsd \tilde M^{\mu\nu}\tilde M^{\rho\sigma*}
 &=  \Hmrd\Hnsd
\left(M^{\mu\nu} + k^\mu p^\nu + p'^\mu k'^\nu \right)
\left(M^{\rho\sigma*} + k^{\rho} p^{\sigma*} + p'^{\rho*}k'^{\sigma} \right)\\
 &= \sum\s{pol.}|M|^2 + \left[(k\cdot p'^*)(k'\cdot p) + \Hc\right].
\end{align}
After all, we have obtained following expression:
\begin{align}
 \sum\s{pol.}|\tilde M|^2 &=
 \sum\s{pol.}|M|^2\qquad \text{(Furthermore $\tilde M=M$)}\notag\\
&= \sum\s{pol.}|\epsilon_\mu^*(k)\epsilon_\nu'^*(k')M^{\mu\nu}|^2
 = \sum\s{pol.}|\epsilon_\mu^*(k)\epsilon_\nu'^*(k')\tilde M^{\mu\nu}|^2\\
&= \Hmrd\Hnsd M^{\mu\nu}M^{\rho\sigma*}\notag\\
&\ne \Hmrd\Hnsd \tilde M^{\mu\nu}\tilde M^{\rho\sigma*}
 = \sum\s{pol.}|\tilde M|^2 + \left[(k\cdot p'^*)(k'\cdot p) + \Hc\right].\notag
\end{align}

\subsection{Phantom Terms in the Gauge Theory}
\label{sec:no-other-term}
You may think we forget to introduce
$\ol\psi\G5\psi$，
$\ol\psi\G5\slashed D\psi$，
$\epsilon^{\mu\nu\rho\sigma}F^a_{\mu\nu}F^a_{\rho\sigma}$，
$\epsilon^{\mu\nu\rho\sigma}D_\mu D_\nu F^a_{\rho\sigma}$
terms, but being a bit careful,
\begin{itemize}
 \item the first two terms are nonsense, for now we use $\PL$ and $\PR$,
 \item the last term is equivalent to the third term as\\[.5zw]\qquad
$\epsilon^{\mu\nu\rho\sigma}D_\mu D_\nu F^a_{\rho\sigma}
=\epsilon^{\mu\nu\rho\sigma}\dfrac12[D_\mu,D_\nu]F^a_{\rho\sigma}
=\dfrac12\epsilon^{\mu\nu\rho\sigma}F^a_{\mu\nu}F^a_{\rho\sigma}$.
\end{itemize}\vspace{.5zw}
Therefore, we have to discuss only the $\epsilon FF$ terms.
If the gauge group is simple, we can take the structure constant as totally antisymmetric, which leads these terms to fall into surface terms as:
\begin{align}
 \epsilon^{\mu\nu\rho\sigma}f^{abc}f^{ade}A^b_\mu A^c_\nu A^d_\rho A^e_\sigma
 &= \epsilon^{\mu\nu\rho\sigma}\left(-f^{acd}f^{abe}-f^{adb}f^{ace}\right)
     A^b_\mu A^c_\nu A^d_\rho A^e_\sigma\notag\\
 &= -2\epsilon^{\mu\nu\rho\sigma}f^{abc}f^{ade}A^b_\mu A^c_\nu A^d_\rho A^e_\sigma\\
 &=  0,\notag
\end{align}
\begin{align}
 \therefore\ \epsilon^{\nu\nu\rho\sigma}F_{\mu\nu}F_{\rho\sigma}
&= 4\epsilon^{\mu\nu\rho\sigma}\Pm A^a_\nu\Pr A^a_\sigma
 + 4g\epsilon^{\mu\nu\rho\sigma}f^{abc}A^a_\mu A^b_\nu\Pr A^c_\sigma\notag\\
&= 2\Pm G^\mu,&
\end{align}
where $G^\mu$ is the Chern--Simons term which is defined as
\begin{equation}
  G^\mu
:= 2\epsilon^{\mu\nu\rho\sigma}
\left( A^a_\nu\Pr A^a_\sigma+\frac13gf^{abc}A^a_\nu A^b_\rho A^c_\sigma \right)
= \epsilon^{\mu\nu\rho\sigma}
\left( A^a_\nu F^a_{\rho\sigma}-\frac13gf^{abc}A^a_\nu A^b_\rho A^c_\sigma \right).
\end{equation}
See Appendix~\ref{sec:cs-instanton} for the instanton effect.

\subsection{Instanton}
\label{sec:cs-instanton}
\TODO{Postpone...}
%%% Local Variables:
%%% TeX-master: "CheatSheet.tex"
%%% End: