### 2019-04-01 spinors.tex
feat: start upgrading the document (in English)

\documentclass[CheatSheet]{subfiles}
\begin{document}

\changefontsizes{10pt}
\section{Spinors}

\changefontsizes{9pt}
\begin{equation}
 (\overline{\psi_1}\psi_2)^* = (\psi_2)^\dagger(\overline\psi_1)^\dagger = \overline{\psi_2}\psi_1.
\end{equation}

\end{document}