### 2019-07-29 mathematics.tex
feat: add a formula for Matrix exponential

\documentclass[CheatSheet]{subfiles}
\begin{document}

\detailstyle
\section{Mathematics}
\subsection{Matrix exponential}
Excerpted from \S2 and \S5 of Hall 2015 \cite{hall2015}:
\begin{alignat}{2}
 &\ee^X := \sum_{m=0}^\infty \frac{X^m}{m!} \text{~~(converges for any $X$)},
\quad&
 &\log X := \sum_{m=1}^\infty (-1)^{m+1}\frac{(A-1)^m}{m} \text{~~(conv.~if $\|A-I\|<1$)}.
\end{alignat}
\begin{alignat}{2}
 &\ee^{\log A} = A \text{~~(if $\|A-I\|<1$)},
\quad&
 &\log \ee^X = X \text{~and~} \|\ee^X-1\| < 1 \text{~~(if $\|X\|<\log2$).}
\end{alignat}
\begin{equation}
 \text{Hilbert-Schmidt norm}:~\|X\|^2 := \sum_{i,j}|X_{ij}|^2 = {\Tr X^\dagger X}.
\end{equation}
Properties:
\begin{alignat*}{4}
 \ee^{(X^\TT)} &= (\ee^X)^\TT,
&\qquad
 \ee^{(X^*)} &= (\ee^X)^*,
&\qquad
 (\ee^X)^{-1} &= \ee^{-X},
&\qquad
 \ee^{YXY^{-1}} &= Y\ee^X Y^{-1},
\end{alignat*}
\begin{alignat*}{3}
 \det \exp X &= \exp\Tr X,
&\qquad
 \frac{\dd}{\dd t}\ee^{tX} &= X\ee^{t X} = \ee^{tX} X
&\qquad
 \ee^{(\alpha+\beta)X} &= \ee^{\alpha X}\ee^{\beta X} \text{~for $\alpha, \beta\in \mathbb C$};
\end{alignat*}
Baker-Campbell-Hausdorff:
\begin{align}
  \ee^X Y \ee^{-X}
&= Y + [X, Y] + \frac1{2!}[X, [X, Y]] + \frac1{3!}[X, [X, [X, Y]]] + \cdots = \ee^{[X,]}Y;
\\
  \ee^X \ee^Y \ee^{-X}
&= \sum_{n=0}^\infty\frac{1}{n!}(\ee^X Y \ee^{-X})^n
 = \exp\left(\ee^{[X,]}Y\right);
\\
  \log(\ee^X \ee^Y)
&= X + \int_0^1\!\dd t\, g(\ee^{[X,}\ee^{t[Y,})Y
\qquad\Bigl[
  g(z) = \frac{\log z}{1-z^{-1}} = 1-\sum_{n=1}^{\infty}\frac{(1-z)^n}{n(n+1)};
\quad
g(\ee^y)=\sum_{n=0}^{\infty}\frac{B_ny^n}{n!}
\Bigr]\\
&= X + Y + \frac12[X,Y] + \frac1{12}[X, [X,Y]] - \frac1{12}[Y, [X,Y]] + \cdots
\quad\text{(Baker-Campbell-Hausdorff).}
\end{align}
\begin{align}
   \log(\ee^X\ee^Y)
&=\sum_{k=1}^\infty \frac{(-1)^{k-1}}{k}\left(\sum_{m,n=0}^{\infty}\frac{X^mY^n}{m!n!}-1\right)^k
&=\sum_{k=1}^\infty
   \sum_{m_1+n_1>0}\cdots\sum_{m_k+n_k>0}\frac{(-1)^{k-1}}{k}
\frac{X^{m_1}Y^{n_1}\cdots X^{m_k}Y^{n_k}}{m_1!n_1!\cdots m_k!n_k!}
\end{align}
\begin{equation}
   \log(\ee^X\ee^Y)=\sum_{k=1}^{\infty}\sum_{m_1+n_1>0}\cdots\sum_{m_k+n_k>0}
\frac{(-1)^{k-1}}{k\sum_{i=1}^{k}(m_i+n_i)}
\frac{
\Bigl([X,\Bigr)^{m_1}
\Bigl([Y,\Bigr)^{n_1}\cdots
\Bigl([X,\Bigr)^{m_k}
\Bigl([Y,\Bigr)^{n_k}]\cdots]
}{m_1!n_1!\cdots m_k!n_k!}
\end{equation}
withe $[X]:=X$ understood.

\end{document}