### 2020-03-18 cosmology.tex
add one cosmology formula

\documentclass[CheatSheet]{subfiles}
\newcommand{\ddP}[2][3]{\frac{\dd^{#1}\vc#2}{(2\pi)^{#1}}}

\begin{document}

\section{Cosmology}
\detailstyle
\subsection{Cosmology formulae}
Assuming that $\sigma$ depends only on $s=(p_A+p_B)^2$, \TODO{(or always true?)}
\begin{align}
& \int\ddP{p_A}\,\,\ddP{p_B}\ee^{-E_A/T}\ee^{-E_B/T}\sigma(s)v_{\text{M\o l}}
\\&=4\int\dd\Pi_A\dd\Pi_B\, \ee^{-(E_A+E_B)/T}\sigma(s)\sqrt{(p_A\cdot p_B)^2-m_A^2m_B^2}
\\&=
\frac{1}{32\pi^4}
\int_{(m_A+m_B)^2}^\infty\dd s
\int_{\sqrt{s}}^\infty\dd E_+\left[\frac{2\sqrt{E_+^2-s}}{s}\mathop{\lambda^{1/2}}(s,m_A^2,m_B^2)\right]
\ee^{-E_+/T}\sigma(s)\left[\frac12\mathop{\lambda^{1/2}}(s,m_A^2,m_B^2)\right]
\\&=
\frac{T}{32\pi^4}
\int_{(m_A+m_B)^2}^\infty\dd s\,s^{3/2} \mathop{K_1}\left(\frac{\sqrt{s}}{T}\right)
\mathop{\lambda}\left(1,\frac{m_A^2}s,\frac{m_B^2}s\right)\sigma(s).
\end{align}

\end{document}