### 2020-08-07 Calculation.tex
write a bit on polarization sum

\documentclass[CheatSheet]{subfiles}

\begin{document}
\summarystyle

\section{Calculation techniques}
\paragraph{Polarization sum}
\begin{alignat}{2}
\text{massless:}\quad
\sum_{\pm}\epsilon^*_\mu(k) \epsilon_\nu(k)
&=
-\eta_{\mu\nu} - \frac{n^2 k_\mu k_\nu}{(n\cdot k)^2} + \frac{n_\mu k_\nu + n_\nu k_\mu}{n\cdot k}
\quad\Bigl(\leadsto -\eta_{\mu\nu}\quad\text{with Ward id.}\Bigr),
\\
\text{massive:}\quad
\sum_{\pm,0}\epsilon^*_\mu(k) \epsilon_\nu(k)
&=
-\eta_{\mu\nu}+\frac{k_\mu k_\nu}{m^2},
\end{alignat}
where $\epsilon\cdot k = 0$ is assumed and $n^\mu$ should satisfy $\epsilon\cdot n=0$, and $k\cdot n\neq 0$ (usually $n^\mu=(1,0,0,0)^\mu$).
\TODO{derivation}
\end{document}